---
title: "TLS data analysis"
author: "Tarini"
date: '2022-06-02'
output: html_document
---

install required packages

```{r}
install.packages("raster")
install.packages("rgdal")
install.packages("sp")
```

install library
```{r}
library(raster)
library(rgdal)
```

Analyze TLS DEM

Upload TLS DEM

```{r}
b = raster("C:/Users/tarin/Documents/santee/aerial_lidar/Extract_Idw_9B.tif")
```

```{r}
plot(b)
```



Distribution of elevation/values
```{r}
hist(b, main="Distribution of elevation values", col="purple")
```


Find 50th percentile elevation value
```{r}
quantile(b)
quantile(b, probs = c(0.70), type=7,names = FALSE)
```


plot elevation upto 70 percentile

```{r}
image(b, zlim=c(0, 0.8102))
```

Histogram with custom breaks
```{r}
hist(b,
     breaks = c(-1, 0, 0.5, 1, 1.5, 2, 3),
     main = "Terrestrial Lidar Histogram with custom breaks",
     xlab = "Elevation (m)",
     ylab = "Number of pixels",
     col = "purple")
```
## Create classification matrix
#-1-0.0 = NA
## 0 to 0.86 = Depressions/channel
## >1 = hummocks

```{r}
reclass_tls <- c(-1, 0.0, NA,
                0.0, 0.86, 1,
                0.86, 0.91, 2,
                0.91, Inf, 3)
reclass_tls
```

# reshape the object into a matrix with columns and rows

```{r}
reclass_tls_m <- matrix(reclass_tls,
                ncol = 3,
                byrow = TRUE)
reclass_tls_m

```
## Reclassify raster
# reclassify the raster using the reclass object - reclass_m


```{r}
microtopo_tls <- reclassify(b, reclass_tls_m)
```


## distribution of pixels as per class
## view reclassified data

```{r}
barplot(microtopo_tls, main = "Number of pixels in each class")
```


## Assign pixels with 0 value as NA
```{r}
microtopo_tls[microtopo_tls == 0] <- NA
```


## Plot reclassified data

```{r}
plot(microtopo_tls, col = c("blue", "green", "purple"),
     main = "Classified microtopography \n depressions, lawns, hummocks")
```

## Lets explore it with Whitebox GAT



























